\bf Model Visualization Techniques for a Social Network Model

### 2017-05-20 paper.Rnw
all plotcon plots in the paper with some description.

\begin{table}
\centering
\begin{tabular}{ll}
\textbf{Structural Effects} \\
outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ \\
reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji}$ \\
transitive triplets  & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\
\textbf{Covariate Effects}\\
covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}z_j$ \\
covariate-ego  & $s_{i5}(x) = z_i\sum_j x_{ij}$ \\
same covariate & $s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(z_i = z_j)$ \\
jumping transitive triplets  & $s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$
\end{tabular}
\caption{\label{tab:effects} Some of the possible effects to be included in the stochastic actor-oriented models in RSiena. There are many more possible effects, but we only consider a select few here. For a complete list, see the RSiena manual \citep{RSiena}.}
\end{table}

@@

<<sienagof, fig.cap="Goodness of fit measures for two network statistics: indegree and outdegree cumulative distributions. On the x-axis is the degree value, and on the y-axis is the number of times that degree value occurs in the data.">>=
load("data/ansnullpaper.rda")
gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship")
gof2 <- sienaGOF(ansnull, IndegreeDistribution, varName ="friendship")
p1 <- plot(gof1)
p2 <- plot(gof2)
grid.arrange(p1, p2)
@

@@

<<alldataviz, fig.cap="The smaller friendship network data we will be modelling throughout the paper.", fig.width=8>>=
actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
                 data.frame(id = 1:16, drink = drink2[,1]),
                 by.x = "X1", by.y = "id", all = T)
actual1$Wave <- 1 
actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))),
                 data.frame(id = 1:16,drink = drink2[,2]),
                 by.x = "X1", by.y = "id", all = T)
actual2$Wave <- 2
actual3 <- merge(data.frame(as.edgelist(as.network(fd2.w3))),
                 data.frame(id = 1:16,drink = drink2[,3]),
                 by.x = "X1", by.y = "id", all = T)
actual3$Wave <- 3
alldat <- rbind(actual1, actual2, actual3)
#alldat$X1 <- paste0("V", alldat$X1)
alldat$X1 <- as.factor(alldat$X1)
#alldat$X2 <- ifelse(!is.na(alldat$X2), paste0("V", alldat$X2), NA)
alldat$X2 <- as.factor(alldat$X2)
write_csv(alldat, "smallfriends4Geomnet.csv")
alldataviz <- ggplot(data = alldat, 
       aes(from_id = X1, to_id = X2, color = as.ordered(drink), group = as.ordered(drink))) +
  geom_net(fiteach = FALSE, directed=TRUE, layout.alg = 'fruchtermanreingold', size = 4, arrowsize = .25, arrowgap = .03, ecolour = 'grey40', labelon=T, labelcolour = 'black', fontsize = 3, vjust = .5) +
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior") + 
  theme_net() +
  facet_wrap(~Wave, labeller = "label_both") + 
  theme(legend.position = 'bottom', panel.background = element_rect(fill = NA, color = "grey30"))
alldataviz
@

@@

# running Siena
#friendship <- sienaDependent(friendshipData)
#alcohol <- varCovar(drink[20:35,])
#mydata <- sienaDataCreate(friendship, alcohol)
#myeffnull <- getEffects(mydata)
#myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:20]

@@

changes <- cumprobs %>% select(ego, alter) %>% unique

@@

@@

@@

We also explore simulations from these different models given parameter values. Using the means from the 1,000 fitted parameter values, we simulated 1,000 observations from each of our three models. We condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models estimated previously. From these simulations, we first create a visualization that represents an ``average" network. To do this, we count occurences of each possible edge in the simulations, resulting in a summary network with weighted edges representing the number of times an edge appeared when simulating from the SAOM.

@@

M1avg$from <- paste0("V", M1avg$from)
M1avg$to <- paste0("V", M1avg$to)
M1avg2$from <- paste0("V", M1avg2$from)
M1avg2$to <- paste0("V", M1avg2$to)
M1avgW2$from <- paste0("V", M1avgW2$from)
M1avgW2$to <- paste0("V", M1avgW2$to)
M2avg$from <- paste0("V", M2avg$from)
M2avg$to <- paste0("V", M2avg$to)
M2avgW2$from <- paste0("V", M2avgW2$from)
M2avgW2$to <- paste0("V", M2avgW2$to)
M2avg2$from <- paste0("V", M2avg2$from)
M2avg2$to <- paste0("V", M2avg2$to)
M3avg$from <- paste0("V", M3avg$from)
M3avg$to <- paste0("V", M3avg$to)
M3avgW2$from <- paste0("V", M3avgW2$from)
M3avgW2$to <- paste0("V", M3avgW2$to)
M3avg2$from <- paste0("V", M3avg2$from)
M3avg2$to <- paste0("V", M3avg2$to)

wave2drink <- drink2[,2]
wave2drink <- as.data.frame(wave2drink)
wave2drink$id <- paste0("V", 1:16)
names(wave2drink)[1] <- "drink"
wave3drink <- drink2[,3]
wave3drink <- as.data.frame(wave3drink)
wave3drink$id <- paste0("V", 1:16)
names(wave3drink)[1] <- "drink"
M1avgcov <- merge(M1avg, wave2drink, by.x = "from", by.y = "id", all = T)
M2avgcov <- merge(M2avg, wave2drink, by.x = "from", by.y = "id", all = T)
M3avgcov <- merge(M3avg, wave2drink, by.x = "from", by.y = "id", all = T)
M1avgW2cov <- merge(M1avgW2, wave3drink, by.x = "from", by.y = "id", all = T)
M2avgW2cov <- merge(M2avgW2, wave3drink, by.x = "from", by.y = "id", all = T)
M3avgW2cov <- merge(M3avgW2, wave3drink, by.x = "from", by.y = "id", all = T)

M1avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM1
rownames(distmatM1) <- distmatM1$from
distmatM1 <- distmatM1[,-1]
distmatM1 <- as.matrix(sqrt(1/distmatM1))
distmatM1[is.na(distmatM1)] <- 0
M1avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM1W2
rownames(distmatM1W2) <- distmatM1W2$from
distmatM1W2 <- distmatM1W2[,-1]
distmatM1W2 <- as.matrix(sqrt(1/distmatM1W2))
distmatM1W2[is.na(distmatM1W2)] <- 0
M2avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM2
rownames(distmatM2) <- distmatM2$from
distmatM2 <- distmatM2[,-1]
distmatM2 <- as.matrix(sqrt(1/distmatM2))
distmatM2[is.na(distmatM2)] <- 0
M2avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM2W2
rownames(distmatM2W2) <- distmatM2W2$from
distmatM2W2 <- distmatM2W2[,-1]
distmatM2W2 <- as.matrix(sqrt(1/distmatM2W2))
distmatM2W2[is.na(distmatM2W2)] <- 0
M3avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM3
rownames(distmatM3) <- distmatM3$from
distmatM3 <- distmatM3[,-1]
distmatM3 <- as.matrix(sqrt(1/distmatM3))
distmatM3[is.na(distmatM3)] <- 0
M3avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM3W2
rownames(distmatM3W2) <- distmatM3W2$from
distmatM3W2 <- distmatM3W2[,-1]
distmatM3W2 <- as.matrix(sqrt(1/distmatM3W2))
distmatM3W2[is.na(distmatM3W2)] <- 0

cols <- brewer.pal(n = 4, "YlOrRd")

plotM1W2 <- ggplot(data = M1avgcov) + 
  geom_net(aes(from_id = from, to_id = to, linewidth = weight/251, color = as.factor(drink)), 
           arrowgap = .03, size = 3,arrowsize = .5,
           directed = TRUE, curvature = .2,
           #layout.alg = "fruchtermanreingold"
           layout.alg = "kamadakawai",
           layout.par = list(elen = distmatM1)
           ) + 
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior") + 
  #labs(title = "Wave 2, M1") + 
  scale_size_continuous(name = "Weight") + 
  theme_net() + 
  theme(plot.title = element_text(hjust = .5), legend.position = 'none',plot.margin=unit(c(0,0,0,0),"mm")) + 
  labs(title = "Model 1 Simulated Wave 2")
plotM1W3 <- ggplot(data = M1avgW2cov) + 
  geom_net(aes(from_id = from, to_id = to, linewidth = weight/388.5, color = as.factor(drink)), 
           arrowgap = .03, size = 3,arrowsize = .5,
           directed = TRUE, curvature = .2,
           #layout.alg = "fruchtermanreingold"
           layout.alg = "kamadakawai",
           layout.par = list(elen = distmatM1W2)
           ) + 
  scale_color_manual(values = cols[2:4], name = "Drinking\nBehavior") + 
  #labs(title = "Wave 3, M1") + 
  scale_size_continuous(name = "Weight") + 
  theme_net() + 
  theme(plot.title = element_text(hjust = .5), legend.position = 'none',plot.margin=unit(c(0,0,0,0),"mm"))
plotM2W2 <- ggplot(data = M2avgcov) + 
  geom_net(aes(from_id = from, to_id = to, linewidth = weight/325, color = as.factor(drink)), 
           arrowgap = .03, size = 3,arrowsize = .5,
           directed = TRUE, curvature = .2,
           #layout.alg = "fruchtermanreingold"
           layout.alg = "kamadakawai",
           layout.par = list(elen = distmatM2)
           ) + 
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior") + 
  #labs(title = "Wave 2, M2") + 
  scale_size_continuous(name = "Weight") + 
  theme_net() + 

>>>>

\begin{table}
\centering
\begin{tabular}{lll}
\textbf{Structural Effects} \\
outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ & \includegraphics[width=3cm]{img/outdegree}\\
reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji}$ & \includegraphics[width=3cm]{img/reciprocity}\\
transitive triplets  & $s_{i3}(x) = \sum_{j,h}& \includegraphics[width=3cm]{img/transtrip} x_{ij}x_{jh}x_{ih}$ \\
\textbf{Covariate Effects}\\
covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}z_j$ & \includegraphics[width=3cm]{img/covaralter} \\
covariate-ego  & $s_{i5}(x) = z_i\sum_j x_{ij}$ & \includegraphics[width=3cm]{img/covearego} \\
same covariate & $s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(z_i = z_j)$ & \includegraphics[width=3cm]{img/covarsim}%\\
%jumping transitive triplets  & $s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$
\end{tabular}
\caption{\label{tab:effects} Some of the possible effects to be included in the stochastic actor-oriented models in RSiena. There are many more possible effects, but we only consider a select few here. For a complete list, see the RSiena manual \citep{RSiena}.}
\end{table}

@@

<<sienagof, fig.cap="Goodness of fit measures for two network statistics: indegree and outdegree cumulative distributions. On the x-axis is the degree value, and on the y-axis is the number of times that degree value occurs in the data.", cache = TRUE, message = FALSE, warning = FALSE>>=
load("data/ansnullpaper.rda")
library(lattice)
gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship")
gof2 <- sienaGOF(ansnull, IndegreeDistribution, varName ="friendship")
p1 <- plot(gof1)
p2 <- plot(gof2)
gridExtra::grid.arrange(p1, p2)
@

@@

<<alldataviz, fig.cap="The smaller friendship network data we will be modelling throughout the paper.", fig.width=8>>=
actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
                 data.frame(id = 1:16, drink = drink2[,1]),
                 by.x = "X1", by.y = "id", all = T)
actual1$Wave <- 1 
actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))),
                 data.frame(id = 1:16,drink = drink2[,2]),
                 by.x = "X1", by.y = "id", all = T)
actual2$Wave <- 2
actual3 <- merge(data.frame(as.edgelist(as.network(fd2.w3))),
                 data.frame(id = 1:16,drink = drink2[,3]),
                 by.x = "X1", by.y = "id", all = T)
actual3$Wave <- 3
alldat <- rbind(actual1, actual2, actual3)
#alldat$X1 <- paste0("V", alldat$X1)
alldat$X1 <- as.factor(alldat$X1)
#alldat$X2 <- ifelse(!is.na(alldat$X2), paste0("V", alldat$X2), NA)
alldat$X2 <- as.factor(alldat$X2)
#write_csv(alldat, "smallfriends4Geomnet.csv")
alldataviz <- ggplot(data = alldat, 
       aes(from_id = X1, to_id = X2, color = as.ordered(drink), group = as.ordered(drink))) +
  geom_net(fiteach = FALSE, directed=TRUE, layout.alg = 'fruchtermanreingold', size = 4, arrowsize = .25, arrowgap = .03, ecolour = 'grey40', labelon=T, labelcolour = 'black', fontsize = 3, vjust = .5) +
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior") + 
  theme_net() +
  facet_wrap(~Wave, labeller = "label_both") + 
  theme(legend.position = 'bottom', panel.background = element_rect(fill = NA, color = "grey30"))
alldataviz
@

@@

# running Siena
#friendship <- sienaDependent(friendshipData)
#alcohol <- varCovar(drink[20:35,])
#mydata <- sienaDataCreate(friendship, alcohol)
#myeffnull <- getEffects(mydata)
#myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  dplyr::select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:20]

@@

changes <- cumprobs %>% dplyr::select(ego, alter) %>% unique

@@

We also wanted to view the complete microstep process from the first wave, which we condition on, to the second wave. The number of steps taken from wave 1 to wave 2 varies. In the simulations from Model 1, the smallest number of steps taken was 58, and the longest was 248, with a mean of 106 and a median of 103. The standard deviation of the 1000 simulations' number of microsteps was 22.8. In Figure~\ref{fig:allsteps}, we see three simulations of the process from wave 1 to wave 2, with wave 1 shown on the left, and wave 2 shown on the right. In each of the three plots, the $y$-axis contains the edges sorted by how often they appear in the networks along the way. We can see that some edges are there in the beginning, but disappear and never come back, while others appear a few steps in, only to dispappear again. There are also some edges that exist in wave 2 that don't appear at all in the microstep process in a given repitition.

<<allsteps, fig.cap="Three repetitions of the microstep process from wave 1 to wave 2. The $x$ axis is the microstep number, with step 0 representing the first wave of data and the final step representing the second wave of data.">>=
smallfriends <- read_csv("data/smallfriends4Geomnet.csv")
names(smallfriends)[1:2] <- c("from", "to")
msall1 <- netvizinf::listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
msall1_df <- plyr::rbind.fill(lapply(X = seq_along(msall1), FUN = function(i, x) {
        x[[i]]$ms <- i - 1
        return(x[[i]])
    }, msall1))
msall1_df <- msall1_df %>% dplyr::select(from, to, ms)
wave2 <- smallfriends %>% filter(Wave == 2) %>% na.omit %>% data.frame()
msall1_df <- msall1_df %>% add_row(from = paste0("V",wave2$from),
                      to = paste0("V",wave2$to), 
                      ms = max(msall1_df$ms)+1)
edges <- msall1_df %>% 
  group_by(from, to) %>%
  summarize(edgeappears = n()) %>% 
  arrange(desc(edgeappears)) %>% ungroup() %>%
  mutate(eid = row_number())

msall1_df2 <- left_join(msall1_df, edges, by = c("from" = "from", "to" = "to"))

p1 <- ggplot(data = msall1_df2) + 
  geom_tile(aes(x = ms, y = -eid, fill = ms)) + 
  geom_vline(xintercept = c(0.5,81.5)) + 
  scale_fill_continuous(low = "red", high = "blue") + 
  theme(legend.position = 'none', axis.ticks.y = element_blank(),
        axis.text.y = element_blank(), axis.title = element_blank()) +
  labs(x = "Microstep")

# rep 2 
msall2 <- netvizinf::listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 2))
msall2_df <- plyr::rbind.fill(lapply(X = seq_along(msall2), FUN = function(i, x) {
        x[[i]]$ms <- i - 1
        return(x[[i]])
    }, msall2))
msall2_df <- msall2_df %>% dplyr::select(from, to, ms)
msall2_df <- msall2_df %>% add_row(from = paste0("V",wave2$from),
                      to = paste0("V",wave2$to), 
                      ms = max(msall2_df$ms)+1)

edges2 <- msall2_df %>% 
  group_by(from, to) %>%
  summarize(edgeappears = n()) %>% 
  arrange(desc(edgeappears)) %>% ungroup() %>%
  mutate(eid = row_number())

msall2_df2 <- left_join(msall2_df, edges2, by = c("from" = "from", "to" = "to"))

p2 <- ggplot(data = msall2_df2) + 
  geom_tile(aes(x = ms, y = -eid, fill = ms)) + 
  geom_vline(xintercept = c(0.5,49.5)) + 
  scale_fill_continuous(low = "red", high = "blue") +
  theme(legend.position = 'none', axis.ticks.y = element_blank(),
        axis.text.y = element_blank(), axis.title = element_blank()) #+
  #labs(x = "Microstep", y = "Edges (most frequent at top)")

# rep 3 
msall3 <- netvizinf::listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 3))
msall3_df <- plyr::rbind.fill(lapply(X = seq_along(msall3), FUN = function(i, x) {
        x[[i]]$ms <- i - 1
        return(x[[i]])
    }, msall3))
msall3_df <- msall3_df %>% dplyr::select(from, to, ms)
msall3_df <- msall3_df %>% add_row(from = paste0("V",wave2$from),
                      to = paste0("V",wave2$to), 
                      ms = max(msall3_df$ms)+1)

edges3 <- msall3_df %>% 
  group_by(from, to) %>%
  summarize(edgeappears = n()) %>% 
  arrange(desc(edgeappears)) %>% ungroup() %>%
  mutate(eid = row_number())

msall3_df2 <- left_join(msall3_df, edges3, by = c("from" = "from", "to" = "to"))

p3 <- ggplot(data = msall3_df2) + 
  geom_tile(aes(x = ms, y = -eid, fill = ms)) + 
  geom_vline(xintercept = c(0.5,106.5)) + 
  scale_fill_continuous(low = "red", high = "blue") +
  theme(legend.position = 'none', axis.ticks.y = element_blank(),
        axis.text.y = element_blank(), axis.title = element_blank()) #+
  #labs(x = "Microstep", y = "Edges (most frequent at top)")

grid.arrange(p1, p2, p3, nrow = 3)
@

@@

Another way to view the model in the data space is through simulation from the model. Since no single simulation from a model is going to look like the data, it's better to visualize many simulations together. For network data, one way to do this is through the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show a summary network created with 1000 simulations from Model 1. Only edges that appear in at least 51 of the 1000 simulations are included in the visualization, with edges that appear more frequently emphasized by thicker linewidths and a bright red color. On either side of the summary network, we show the actual data, wave 1 on the left, and wave 2 on the right. Comparing the three visualizations side-by-side shows us that the summary network of simulations of wave 2 looks more like the true wave 2 than wave 1, which is a good indication that the model is capturing the true data structure. It's not ideal, however, because there appear to be too many reciprocated ties in the summary network, far more than there are in the data. This could indicate that we should remove the reciprocity parameter from the model.

<<summnetM1>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- s50a[20:35,1]
drink2 <- s50a[20:35,2]
drink3 <- s50a[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
M1sims1000 <- saom_simulate(dat = mydata, 
                            struct = M1eff,
                       parms = as.numeric(M1parms), 
                       N = 1000)
M2sims1000 <- saom_simulate(dat = mydata, 
                            struct = M2eff,
                       parms = as.numeric(M2parms), 
                       N = 1000)
M3sims1000 <- saom_simulate(dat = mydata, 
                            struct = M3eff,
                       parms = as.numeric(M3parms), 
                       N = 1000)
M1simsdf <- sims_to_df(M1sims1000)
M2simsdf <- sims_to_df(M2sims1000)
M3simsdf <- sims_to_df(M3sims1000)
M1avgW2 <- M1simsdf %>% filter(!is.na(to) & wave == 2) %>%
  group_by(from, to) %>% 
  summarise(count = n()) %>% 
  mutate(weight = ifelse(from == to, 0, count))
M2avgW2 <- M2simsdf %>% filter(!is.na(to) & wave == 2) %>%
  group_by(from, to) %>% 
  summarise(count = n()) %>% 
  mutate(weight = ifelse(from == to, 0, count))
M3avgW2 <- M3simsdf %>% filter(!is.na(to) & wave == 2) %>%
  group_by(from, to) %>% 
  summarise(count = n()) %>% 
  mutate(weight = ifelse(from == to, 0, count))

# make a df of wave 2, wave 1, and the three averages and facet. 
names(actual1)[1:2] <- c("from", "to")
names(actual2)[1:2] <- c("from", "to")
actual1$count <- 1
actual1$weight <- 1
actual2$count <- 1
actual2$weight <- 1
actual1 <- actual1 %>% dplyr::select(from,to, count, weight)
actual2 <- actual2 %>% dplyr::select(from,to, count, weight)
actual1$cat <- "First Observation"
actual2$cat <- "Second Observation"

avgW2M1 <- M1avgW2 %>% ungroup() %>% 
  filter(weight > 50) %>% 
  mutate(from = as.factor(from), to = as.factor(to),
         cat = "Model1")
add1 <- setdiff(as.character(1:16), unique(c(as.character(avgW2M1$from), as.character(avgW2M1$to)))) 
avgW2M1 %>% add_row(from = add1, to = NA, count = NA, weight = NA, cat = "Model1") -> avgW2M1
avgW2M2 <- M2avgW2 %>% ungroup() %>% 
  filter(weight > 50) %>% 
  mutate(from = as.factor(from), to = as.factor(to),
         cat = "Model2")
add2 <- setdiff(as.character(1:16), unique(c(as.character(avgW2M2$from), as.character(avgW2M2$to)))) 
avgW2M2 %>% add_row(from = add2, to = NA, count = NA, weight = NA, cat = "Model2") -> avgW2M2
avgW2M3 <- M3avgW2 %>% ungroup() %>% 
  filter(weight > 50) %>% 
  mutate(from = as.factor(from), to = as.factor(to),
         cat = "Model3")
add3 <- setdiff(as.character(1:16), unique(c(as.character(avgW2M3$from), as.character(avgW2M3$to)))) 
avgW2M3 %>% add_row(from = add3, to = NA, count = NA, weight = NA, cat = "Model3") -> avgW2M3

combinedavgs <- rbind(actual1, actual2, avgW2M1, avgW2M2, avgW2M3)
combinedavgs %>% group_by(cat) %>% 
  mutate(linewidth = weight / max(weight,na.rm = T)) -> combinedavgs

combinedavgs %>% filter(cat == "Model1") %>% mutate(logweight = log(weight)) -> t1 
colors <- tweenr::tween_color(data = c("#969696", "#67000d"), n = 26, ease = 'linear')
t1 %>% arrange(logweight) -> t1
t1$color <- NA
t1$color[1:26] <- colors[[1]]

p1 <- ggplot(data = t1) + 
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.1, "inches")), ecolour = t1$color,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 3, arrowgap = .02,
           singletons = T) +
  theme_net() + labs(title = "Summary Network: M1") + theme(plot.background = element_rect(color = 'black'))
p2 <- ggplot(data = filter(combinedavgs, cat == "First Observation")) + 
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.1, "inches")), ecolour = "#969696",
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 3, arrowgap = .02,
           singletons = T) +
  theme_net() + labs(title = "Wave 1") + theme(plot.background = element_rect(color = 'black'))
p3 <- ggplot(data = filter(combinedavgs, cat == "Second Observation")) + 
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.1, "inches")), ecolour = "#969696",
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 3, arrowgap = .02,
           singletons = T) +
  theme_net() + labs(title = "Wave 2")  + theme(plot.background = element_rect(color = 'black'))
grid.arrange(p2, p1, p3, nrow = 1)
@

@@

We also explore simulations from these different models given parameter values. Using the means from the 1,000 fitted parameter values, we simulated 1,000 observations from each of our three models. We condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models estimated previously. From these simulations, we first create a visualization that represents an ``average" network. To do this, we count occurences of each possible edge in the simulations, resulting in a summary network with weighted edges representing the number of times an edge appeared in the simulated wave 2 when simulating from the SAOM 1000 times. In Figure~\ref{averagenet}, we show the first wave, the ``average'' network from the three models we fit, and the second wave. Comparing the three averages to waves 1 and 2, we see that they have very similar structure to wave 2. Model 2, which included the transitive triplet parameter, seems to have created more group strucutre overall than models 1 and 3. This could be an indication that, despite the highly significant nature of the parameter $\beta_3$ according to the $t$-tests in \verbatim{RSiena}, the transitive triplet parameter should not be included because it does not result in a structure similar to the data.

@@

# M1avg$from <- paste0("V", M1avg$from)
# M1avg$to <- paste0("V", M1avg$to)
# M1avg2$from <- paste0("V", M1avg2$from)
# M1avg2$to <- paste0("V", M1avg2$to)
# M1avgW2$from <- paste0("V", M1avgW2$from)
# M1avgW2$to <- paste0("V", M1avgW2$to)
# M2avg$from <- paste0("V", M2avg$from)
# M2avg$to <- paste0("V", M2avg$to)
# M2avgW2$from <- paste0("V", M2avgW2$from)
# M2avgW2$to <- paste0("V", M2avgW2$to)
# M2avg2$from <- paste0("V", M2avg2$from)
# M2avg2$to <- paste0("V", M2avg2$to)
# M3avg$from <- paste0("V", M3avg$from)
# M3avg$to <- paste0("V", M3avg$to)
# M3avgW2$from <- paste0("V", M3avgW2$from)
# M3avgW2$to <- paste0("V", M3avgW2$to)
# M3avg2$from <- paste0("V", M3avg2$from)
# M3avg2$to <- paste0("V", M3avg2$to)

ggplot(data = combinedavgs) + 
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.1, "inches")), ecolour = "#969696", fiteach = T, 
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 3, arrowgap = .02,
           singletons = F) + 
  theme_net() + theme(panel.background = element_rect(color = 'black')) + 
  facet_wrap(~cat)

# wave2drink <- drink2[,2]
# wave2drink <- as.data.frame(wave2drink)
# wave2drink$id <- paste0("V", 1:16)
# names(wave2drink)[1] <- "drink"
# wave3drink <- drink2[,3]
# wave3drink <- as.data.frame(wave3drink)
# wave3drink$id <- paste0("V", 1:16)
# names(wave3drink)[1] <- "drink"
# M1avgcov <- merge(M1avg, wave2drink, by.x = "from", by.y = "id", all = T)
# M2avgcov <- merge(M2avg, wave2drink, by.x = "from", by.y = "id", all = T)
# M3avgcov <- merge(M3avg, wave2drink, by.x = "from", by.y = "id", all = T)
# M1avgW2cov <- merge(M1avgW2, wave3drink, by.x = "from", by.y = "id", all = T)
# M2avgW2cov <- merge(M2avgW2, wave3drink, by.x = "from", by.y = "id", all = T)
# M3avgW2cov <- merge(M3avgW2, wave3drink, by.x = "from", by.y = "id", all = T)
# 
# M1avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM1
# rownames(distmatM1) <- distmatM1$from
# distmatM1 <- distmatM1[,-1]
# distmatM1 <- as.matrix(sqrt(1/distmatM1))
# distmatM1[is.na(distmatM1)] <- 0
# M1avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM1W2
# rownames(distmatM1W2) <- distmatM1W2$from
# distmatM1W2 <- distmatM1W2[,-1]
# distmatM1W2 <- as.matrix(sqrt(1/distmatM1W2))
# distmatM1W2[is.na(distmatM1W2)] <- 0
# M2avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM2
# rownames(distmatM2) <- distmatM2$from
# distmatM2 <- distmatM2[,-1]
# distmatM2 <- as.matrix(sqrt(1/distmatM2))
# distmatM2[is.na(distmatM2)] <- 0
# M2avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM2W2
# rownames(distmatM2W2) <- distmatM2W2$from
# distmatM2W2 <- distmatM2W2[,-1]
# distmatM2W2 <- as.matrix(sqrt(1/distmatM2W2))
# distmatM2W2[is.na(distmatM2W2)] <- 0
# M3avg[,-3] %>% spread(to, weight) %>% data.frame-> distmatM3
# rownames(distmatM3) <- distmatM3$from
# distmatM3 <- distmatM3[,-1]
# distmatM3 <- as.matrix(sqrt(1/distmatM3))
# distmatM3[is.na(distmatM3)] <- 0
# M3avgW2[,-3] %>% spread(to, weight) %>% data.frame-> distmatM3W2
# rownames(distmatM3W2) <- distmatM3W2$from
# distmatM3W2 <- distmatM3W2[,-1]
# distmatM3W2 <- as.matrix(sqrt(1/distmatM3W2))
# distmatM3W2[is.na(distmatM3W2)] <- 0

# cols <- brewer.pal(n = 4, "YlOrRd")
# 
# plotM1W2 <- ggplot(data = M1avgcov) + 
#   geom_net(aes(from_id = from, to_id = to, linewidth = weight/251, color = as.factor(drink)), 
#            arrowgap = .03, size = 3,arrowsize = .5,
#            directed = TRUE, curvature = .2,
#            #layout.alg = "fruchtermanreingold"
#            layout.alg = "kamadakawai",
#            layout.par = list(elen = distmatM1)
#            ) + 
#   scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior") + 
#   #labs(title = "Wave 2, M1") + 
#   scale_size_continuous(name = "Weight") + 
#   theme_net() + 
#   theme(plot.title = element_text(hjust = .5), legend.position = 'none',plot.margin=unit(c(0,0,0,0),"mm")) + 
#   labs(title = "Model 1 Simulated Wave 2")
# plotM1W3 <- ggplot(data = M1avgW2cov) + 
#   geom_net(aes(from_id = from, to_id = to, linewidth = weight/388.5, color = as.factor(drink)), 
#            arrowgap = .03, size = 3,arrowsize = .5,
#            directed = TRUE, curvature = .2,
#            #layout.alg = "fruchtermanreingold"
#            layout.alg = "kamadakawai",
#            layout.par = list(elen = distmatM1W2)
#            ) + 
#   scale_color_manual(values = cols[2:4], name = "Drinking\nBehavior") + 
#   #labs(title = "Wave 3, M1") + 
#   scale_size_continuous(name = "Weight") + 
#   theme_net() + 
